---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# discretes

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/probaverse/discretes/graph/badge.svg)](https://app.codecov.io/gh/probaverse/discretes)
[![R-CMD-check](https://github.com/probaverse/discretes/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/probaverse/discretes/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The discretes package provides a framework for representing numeric series that may be finite or infinite. Think `1:Inf`, without storing all the values explicitly.

Series can be traversed, combined using arithmetic operations, tested for membership, and queried for limit points ("sinks"), without explicit enumeration of all elements.

The name "discretes" reflects the original use case of representing the supports of discrete probability distributions, which are often infinite but enumerable.
The package is not limited to probability applications: it is designed as a general-purpose tool for working with enumerable numeric series.


## Installation

Install discretes from CRAN with:

``` r
install.packages("discretes")
```

## Examples

```{r}
library(discretes)
```

While vectors in R must be finite, this is not a problem for discretes.

```{r}
natural0()
```

What's the next discrete value after 10 in the series of integers? Or the previous 5 values from 1.3?

```{r}
next_discrete(integers(), from = 10)
prev_discrete(integers(), from = 1.3, n = 5)
```

Test whether values are discrete points in a series.

```{r}
has_discretes(natural1(), c(0, 1))
```

Perform arithmetic operations on series.

```{r}
1 / 2^integers()
```

Arithmetic operations on series return new series.
See the [transformations]() vignette to learn more about what's allowed (hint: non-monotonic functions are not), and how to use transformations that are not supported out of the box.

### Sinks

That last series above, `1 / 2^integers()`, has a sink[^1] at 0 (approached from the right) and a sink at infinity, best seen by plotting:

[^1]: Or more precisely, a limit point.

```{r}
x <- 1 / 2^integers()
plot(x)
```

Notice that there are infinitely many discrete values close to 0.

```{r}
num_discretes(x, from = 0, to = 1)
```

There is no such thing as a "next" value when looking left of the sink.

```{r}
next_discrete(x, from = -1)
```

You can ask whether a sink exists directly.

```{r}
has_sink(x, from = 0, to = 1)
```


## Limitations

The series supported by the package include arithmetic series like integers, finite series from a numeric vector, and some [transformations]() of those series.
Specialized series like the Fibonacci numbers are not explicitly supported, but could be added by a user.
Dense countable sets like the rational numbers are not supported because they do not have a well-defined notion of local successor/predecessor.

## Similar Packages

- ['Zseq'](https://cran.r-project.org/package=Zseq) provides access to named integer sequences (e.g., Fibonacci numbers, prime numbers), but does not provide a general framework for constructing and transforming numeric series.
-	['sets'](https://cran.r-project.org/web/packages/sets/index.html) focuses on finite set operations and abstract set algebra, rather than structured numeric series.
-	['set6'](https://CRAN.R-project.org/package=set6) supported infinite sets via object-oriented abstractions, but is no longer available on CRAN.
- ['peruse'](https://jacgoldsm.github.io/peruse/) provides tools for iterating general sequences, but does not focus on algebraic manipulation or structured numeric series.

## Acknowledgements

Development of this package would not have been possible without the funding and support of [BGC Engineering Inc.](https://www.bgcengineering.ca/), the [European Space Agency](https://www.esa.int/), and the [Politecnico di Milano](https://www.polimi.it/).
The need for this package arose from work on the [probaverse](https://probaverse.com/) project, which aims to provide tools for probabilistic modeling and inference in R.