---
title: "Creating numeric series"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating numeric series}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(discretes)
```

The discretes package lets you create numeric series in two main ways:

**Arithmetic series** — Use `arithmetic()` or the wrappers `integers()`, `natural1()`, and `natural0()` for common cases (all integers, positive integers starting at 1, non-negative integers starting at 0).

**From a numeric vector** — Use `as_discretes()` to turn an existing numeric vector into a numeric series.

Once you have a base series, you can create new ones by subsetting, combining, or transforming.

## Subsetting and combining

- **Subsetting**: `dsct_keep()` and `dsct_drop()` restrict a series to or outside of an interval.
- **Combining**: `dsct_union()` merges multiple series into one.

## Arithmetic and standard functions

**Arithmetic** — The operations `+`, `-`, `*`, `/`, `^` are supported when one side is a number and the other is a numeric series. For example:

```{r}
integers()^2
2 * natural1() - 1   # odd positive integers
1 / 2^integers()     # reciprocals of powers of 2
```

Length-0 vectors are allowed, too, but result in empty series:

```{r}
numeric(0L) * natural1()
```

**Mathematical functions** — `exp()` is supported. For non-negative series, `log()`, `log2()`, `log10()`, and `sqrt()` are also supported:

```{r}
log(natural0())
```

Other functions may be able to be applied as well; use `dsct_transform()` for those (see below).

## Custom transformations: `dsct_transform()`

For other transformations, use `dsct_transform()`. On a numeric vector, the function is applied directly:

```{r}
dsct_transform(0:10, cos) # Same as cos(0:10)
```

When transforming an object of class `"discretes"` there are more restrictions so that querying the series can be defined
from the base series.

1. The function and its inverse must be vectorized and element-wise (no `cumsum()`-style dependence on other elements).
2. You must supply the **domain** and **range** of the function. This defaults to all real numbers.
3. The function must be [strictly monotonic](https://en.wikipedia.org/wiki/Monotonic_function) over the given domain.

Here is an example of applying the function `pnorm()`. The range is (0, 1), so it must be given explicitly:

```{r}
dsct_transform(
  natural0(),
  fun = pnorm,
  inv = qnorm,
  range = c(0, 1)
)
```

For a function that is monotonic only on an interval, restrict the domain.
For example, `cos()` is decreasing on `[0, pi]` with range `[-1, 1]`, and inverse `acos()`.
Specify `dir = "decreasing"` to indicate that this is a decreasing function.

```{r}
dsct_transform(
  0:3,
  fun = cos,
  inv = acos,
  domain = c(0, pi),
  range = c(-1, 1),
  dir = "decreasing"
)
```
