---
title: "Signed Zero"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{signed_zero}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(discretes)
```

In R, zero has a latent sign: `+0` and `-0` print the same but behave differently under reciprocation:
`1 / 0` is `Inf` and `1 / -0` is `-Inf`.
When a numeric series contains zero, that sign can matter for derived series (for example, after inversion).

For numeric series, rules around signed zero follow those of base R.

The functions `has_negative_zero()` and `has_positive_zero()` report whether `-0` or `+0` is a discrete value in the series.
For example, the integers contain a single positive zero:

```{r}
has_negative_zero(integers())
has_positive_zero(integers())
```

Negation can induce negative zeroes (although note that, in R, -0 is only possible as type "double" and not of type "integer").

```{r}
has_negative_zero(-1.5 * integers())
```

Like numeric vectors, a series can contain both signs of zero.
Only one zero is enumerated by `next_discrete()`, `prev_discrete()`, or `get_discretes_*()`.

```{r}
x <- dsct_union(-0, 0)
x
has_positive_zero(x)
has_negative_zero(x)
```

But, both signs remain latent and appear when the series is inverted:

```{r}
1 / x
```

When choosing to pull positive or negative zero from the series:

- For numeric vectors, behaviour is delegated to `base::unique()`.
- For combined series through `dsct_union()`, the zero from the first series that has a zero is taken.
- Otherwise, positive zero is selected.
