---
title: "tolerance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tolerance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(discretes)
```

## What is `tol`?

Many operations in **discretes** involve comparing a *queried* numeric value to
the discrete values in a numeric series. Because most decimal fractions are not
represented exactly in floating point (e.g. `0.1`), values that "should" match
can differ by a tiny amount after arithmetic.

In this package, `tol` is an **absolute tolerance** used when deciding whether a
query value is "close enough" to a discrete value in the series.

- Setting **`tol = 0`** requests **exact** comparisons.
- The default is `sqrt(.Machine$double.eps)`, which is small but usually enough
  to ignore round-off noise.

## Where is `tol` actually used?

Conceptually, `tol` is only used at the points where the package must make a
yes/no decision about membership or boundaries. Most other functions accept
`tol` and simply pass it along to wherever that decision is made.

The two most important "leaf" cases are:

- **Arithmetic progressions** (`arithmetic()`): membership is checked by testing
  whether the implied index \((x - representative) / spacing\) is *integer-ish*.
- **Explicit numeric sets** (plain numeric vectors, or `as_discretes(numeric)`):
  membership is checked by testing whether a queried value is within `tol` of
  any stored value.

Transformations (like `exp(as_discretes(...))`) work by mapping queried points
back to the root space and then applying the same membership logic there, using
the same `tol`.

## Example: `tol` with `arithmetic()`

Without a tolerance, values can be hard to recognize as discrete values in an
arithmetic series because the implied index might be `2.999999999...` instead of `3`.

```{r}
x <- arithmetic(representative = 0, spacing = 0.1)
has_discretes(x, 0.3)
has_discretes(x, 0.3, tol = 0)
```

## Example: `tol` with an explicit numeric set

Even if a set is represented as a numeric vector, you can still run into tiny
floating-point mismatches.

```{r}
v <- c(0, 0.1, 0.2, 0.1 * 3)   # last entry is not exactly 0.3

has_discretes(v, 0.3)
has_discretes(v, 0.3, tol = 0)
```

If you want "mathematical set" behaviour for an explicit set, use `tol = 0`.
If you want "numerical computing" behaviour (robust to round-off), keep `tol`
positive.

## Example: `tol` propagates through transformations

When you transform a numeric series, membership checks happen in the *root space*.
This means `tol` is applied *before* the transformation (via the inverse map),
not directly on the transformed discrete values. In practice, this can make
membership checks behave differently than they would for an ordinary numeric vector.

```{r}
raw <- 10^(-10:10)
frozen <- 10^as_discretes(-10:10)

# A realistic way to get a value that is "almost" 5: repeated addition.
exponent <- tail(cumsum(rep(0.1, 5000)), 1) / 100
q <- 10^exponent

has_discretes(raw, q)
has_discretes(frozen, q)
has_discretes(frozen, q, tol = 0)
```

## Choosing a `tol`

- **Default is usually fine**: `sqrt(.Machine$double.eps)` is intended to smooth
  over tiny floating-point noise.
- **Use `tol = 0` for exactness**: when you truly want explicit sets to behave
  exactly (e.g., when values were constructed to be exactly representable).
- **Increase `tol` for noisy values**: if your values come from computation or
  measurement and you want a wider "snap-to" zone, increase `tol` accordingly.
