---
title: "Tolerance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tolerance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(discretes)
```


Many operations in the discretes package involve comparing a *queried* numeric value to
the discrete values in a numeric series. Because most decimal fractions are not
represented exactly in floating point, values that "should" match
can differ by a tiny amount after arithmetic.

In this package, `tol` is an **absolute tolerance** used when deciding whether a
query value is "close enough" to a discrete value in the series.

- Setting **`tol = 0`** requests **exact** comparisons.
- The default is `sqrt(.Machine$double.eps)`, which is small but usually enough
  to ignore round-off noise.

## Where is `tol` actually used?

Tolerance is not applied to the series you see at the top level. 
It is **passed down** through any nested or transformed series and used **only when membership is actually tested** at the underlying series.

The two places where it is used are:

- **Arithmetic progressions** (`arithmetic()`): membership is decided by the
  implied step index \((x - representative) / spacing\). Because of floating
  point, this index is rarely an exact integer (e.g. `(0.3 - 0) / 0.1`). A
  value is treated as a discrete value if that index is within `tol` of an
  integer.
- **Numeric-vector-based series** (plain numeric vectors or `as_discretes()`):
  membership is decided by testing whether the queried value is within `tol` of
  any stored discrete value.

For transformed series (e.g. `1 / arithmetic()`), the queried value is mapped
back to the underlying series and the same logic is applied there with the same
`tol`; no separate tolerance is applied to the transformed values.

## Example: `tol` with `arithmetic()`

Without a tolerance, values can be hard to recognize as discrete values in an
arithmetic series because the implied index might be `2.999999999` instead of `3`.

```{r}
x <- arithmetic(representative = 0, spacing = 0.1)
has_discretes(x, 0.3)
has_discretes(x, 0.3, tol = 0)
```

## Example: `tol` with an explicit numeric series

Even if a series is represented as a numeric vector, you can still run into tiny
floating-point mismatches.

```{r}
v <- c(0, 0.1, 0.2, 0.1 * 3)   # last entry is not exactly 0.3

has_discretes(v, 0.3)
has_discretes(v, 0.3, tol = 0)
```

If you want "mathematical series" behaviour for an explicit series, use `tol = 0`.
If you want "numerical computing" behaviour (robust to round-off), keep `tol` positive.

## Example: preserving numeric vectors

When a numeric vector is converted to a "discretes" object via `as_discretes()`, the root numeric vector is preserved, even if the series gets transformed.

Consider transforming a numeric vector vs. that same vector expressed as a "discretes" object.

```{r}
raw <- 10^(-5:5)
raw
preserved <- 10^as_discretes(-5:5)
preserved
```

The numeric vector gets transformed right away, whereas it remains as the base series when transformed as a "discretes" object. This has implications because errors are propagated differently.

For example, an exponent slightly off by `1e-10` results in a more significant error after transformation that may not be within the default `tol`:

```{r}
# Exponent slightly off from `5`
q <- 10^(5 - 1e-10)
# Error after transformation
10^5 - q
```

This means that membership on the eagerly transformed numeric vector fails, but not so when the vector is preserved.
This is because tolerance is applied to the root series.

```{r}
has_discretes(raw, q)
has_discretes(preserved, q)
```

## Choosing a `tol`

- **Default is usually fine**: `sqrt(.Machine$double.eps)` is intended to smooth over tiny floating-point noise.
- **Use `tol = 0` for exactness**: when you truly want explicit sets to behave exactly, perhaps useful in some situations when working with 'integer' storage mode.
- **Increase `tol` for noisy values**: if your values come from computation or measurement and you are noticing unusual membership failures, increase `tol` accordingly.
