---
title: "transformations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{transformations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(discretes)
```

Here is a list of all transformations that are currently implemented.

Operations `+`, `-`, `*`, `/`, `^` are supported when a number is on one side, and a numeric series is on the other.

Operations that are currently not supported:

- Operations between two numeric series.
- Operations that are not strictly increasing / decreasing within the range of the series.

For example, if a series has both negativea and positive values, it can be cubed, but squaring is not supported because it is not strictly monotonic.

```{r}
integers()^3
```

```{r, error=TRUE}
integers()^2
```

But squaring is supported if the series is only positive or only negative.

```{r}
natural0()^2
```

```{r}
(-natural0())^2
```

The following mathematical functions are supported:

- `sqrt()`, on series with non-negative values.
- `exp()`
- `log()`, `log2()`, and `log10()` are supported on series with non-negative values.

## Other Transformations

If there are other transformations you'd like to apply, you can use `dsct_transform()`. Make sure:

1. The function is vectorized and operates element-wise; likewise for its inverse. This means functions like `cumsum()` that depend on other values of the vector are not supported.
2. The function is strictly increasing over the intended domain. If you'd like to apply a decreasing function, consider applying the negative of the function, and then applying the negative of the inverse function to transform back.

Let's say you'd like to apply the `pnorm()` function to a series. You can't do `pnorm(integers())`, but you can instead put it in `dsct_transform()`. The range of the function is not the set of real numbers, so we need to specify it explicitly.

```{r}
dsct_transform(
  natural0(),
  fun = pnorm,
  inv = qnorm,
  range = c(0, 1)
)
```

You can also use functions that are not increasing over the entire real line, but are increasing over a specified domain. For example `sin()` is increasing between `-pi/2` and `pi/2`, evaluating to between -1 and 1, with inverse `asin()`. 

```{r}
dsct_transform(
  -3:3,
  fun = sin,
  inv = asin,
  domain = c(-pi/2, pi/2),
  range = c(-1, 1)
)
```

